---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# **Status of coral reefs of the world: 2025**


## 1. Introduction

## 2. Code

### A. Cleaning and selection

* `a1_extract-reef-aca.js` Export [Allen Coral Atlas](https://developers.google.com/earth-engine/datasets/catalog/ACA_reef_habitat_v2_0) coral reef distribution data as .shp at a 500 m resolution from Google Earth Engine (GEE)
* `a2_clean-reefs.R` Add GCRMN region to coral reef distribution and create 100 km coral reef buffer

### B. Extraction

### C. Models

### D. Figures and tables

### E. Functions

* `data_descriptors.R` Get number of sites, surveys, datasets, first and last year of monitoring.
* `graphical_par.R` Graphical parameters, including colors and fonts.
* `theme_graph.R` Main ggplot theme for the plots of the reports.
* `theme_map.R` Main ggplot theme for the maps of the reports.

## 3. Reproducibility parameters

```{r echo=FALSE, message=FALSE}

# 1. Print session info ----

devtools::session_info(c("sf", "tidyverse", "tidymodels"))

```
