---
title: "Chapter `r nb_chapter`"
subtitle: "`r region_i`"
output:
  word_document:
    reference_docx: styles.docx
editor: source
---

```{r echo=FALSE, message=FALSE}

require(tidyverse)
require(readxl)
require(flextable)

fig_counter <- 1

```

\

\
<div custom-style="textb">

**Authors**: Firstname, Surname^1^,  Firstname, Surname^2^

\

^1^ Affiliation name
\
^2^ Affiliation name

</div>

\

\

\

<div custom-style="text">

**Citation**: Firstname, S.,  Firstname, S., *et al.* (2026). *Chapter `r nb_chapter` - Status and Trends of Coral Reefs in `r region_i`*. In Dallison T., Wicquart J., Gonzalez-Rivero, M., Towle, E.K., Brigdale, A., Staub, F., Planes, S. *Status and Trends of Coral Reefs of the World: 2025*. Global Coral Reef Monitoring Network (GCRMN) and International Coral Reef Initiative (ICRI). DOI.

</div>

\newpage

## Geographic information and context

<div custom-style="explanations">
**[Number of words of the section: 500]**
</div>

<div custom-style="text">

```{r echo=FALSE, message=FALSE}

data_reef_extent <- read.csv("../../figs/08_text-gen/reefs_extent.csv") %>% 
  filter(region == region_i)

extent_rank <- data_reef_extent %>% 
  filter(subregion == "All") %>% 
  select(pos_world) %>% 
  mutate(pos_world = as.character(pos_world),
         pos_world = case_when(pos_world == "1" ~ "1^st^",
                               pos_world == "2" ~ "2^nd^",
                               pos_world == "3" ~ "3^rd^",
                               TRUE ~ paste0(pos_world, "^th^"))) %>% pull()

extent_abs <- data_reef_extent %>% 
  filter(subregion == "All") %>% 
  select(reef_extent_abs) %>% 
  mutate(reef_extent_abs = format(round(reef_extent_abs, 0),
                                  big.mark = ",", scientific = FALSE)) %>% pull()

extent_rel <- data_reef_extent %>% 
  filter(subregion == "All") %>% 
  select(reef_extent_rel_world) %>% 
  mutate(reef_extent_rel_world = format(round(reef_extent_rel_world, 1))) %>% pull()

axis_lat <- read.csv("../../figs/08_text-gen/region_characteristics.csv") %>% 
  filter(region == region_i) %>% 
  select(lat_axis) %>% 
  mutate(lat_axis = format(lat_axis, big.mark = ",", scientific = FALSE)) %>% 
  pull()

axis_long <- read.csv("../../figs/08_text-gen/region_characteristics.csv") %>% 
  filter(region == region_i) %>% 
  select(long_axis) %>% 
  mutate(long_axis = format(long_axis, big.mark = ",", scientific = FALSE)) %>% 
  pull()

nb_countries <- read.csv("../../figs/08_text-gen/region_characteristics.csv") %>% 
  filter(region == region_i) %>% 
  select(nb_countries) %>% 
  pull()

border_north <- read.csv("../../figs/08_text-gen/region_characteristics.csv") %>% 
  filter(region == region_i) %>% 
  select(border_north) %>% 
  pull()

border_south <- read.csv("../../figs/08_text-gen/region_characteristics.csv") %>% 
  filter(region == region_i) %>% 
  select(border_south) %>% 
  pull()

border_east <- read.csv("../../figs/08_text-gen/region_characteristics.csv") %>% 
  filter(region == region_i) %>% 
  select(border_east) %>% 
  pull()

border_west <- read.csv("../../figs/08_text-gen/region_characteristics.csv") %>% 
  filter(region == region_i) %>% 
  select(border_west) %>% 
  pull()

denomination <- read.csv("../../figs/08_text-gen/region_characteristics.csv") %>% 
  filter(region == region_i) %>% 
  select(denomination) %>% 
  pull()

```

The `r region_i` GCRMN region ranges over a longitudinal axis of `r axis_long` km and a latitudinal axis of `r axis_lat` km. The region hosts `r extent_abs` km^2^ of coral reefs, which represent `r extent_rel`% of the world’s coral reef extent, and `r region_i` is the `r extent_rank` largest GCRMN region in terms of coral reef extent. Coral reefs reside within the economic exclusive zones of `r nb_countries` `r denomination` within the region, ranging from `r border_north` in the North to `r border_south` in the South, and from `r border_west` in the West to `r border_east` in the East (**Fig. 2.`r nb_chapter`.1**).

```{r echo=FALSE, message=FALSE}

figure_path <- paste0("../../figs/03_part-2/fig-1/", str_to_lower(region_name), ".png")

```

![](`r figure_path`){width=100%}

**Figure 2.`r nb_chapter`.`r fig_counter`.** The distribution of coral reefs within the `r region_i` region and its respective sub-regions. 
White polygons and numbers represent sub-regions. Coral reefs are delineated with a 20 km buffer to enhance interpretation, and only coral reefs within the region are represented. The scale bar provides distance at the equator and may not accurately represent distance over the entire latitudinal range of the region. The borders and nomenclature represented on this map are not endorsed by the GCRMN.

\newpage

```{r echo=FALSE, message=FALSE}

nb_subregions <- data_reef_extent %>% 
  filter(subregion != "All") %>% 
  select(subregion) %>% 
  summarise(nb_subregions = length(subregion)) %>% 
  pull()

max_reef_extent_nb <- data_reef_extent %>% 
  filter(subregion != "All") %>% 
  filter(reef_extent_abs == max(reef_extent_abs)) %>% 
  mutate(subregion = str_trim(gsub("[A-z]", "", subregion))) %>% select(subregion) %>% pull()
  
max_reef_extent_value_abs <- data_reef_extent %>% 
  filter(subregion != "All") %>% 
  filter(reef_extent_abs == max(reef_extent_abs)) %>% 
  select(reef_extent_abs) %>% 
  mutate(reef_extent_abs = format(round(reef_extent_abs, 0),
                                  big.mark = ",", scientific = FALSE)) %>% pull()

max_reef_extent_value_rel <- data_reef_extent %>% 
  filter(subregion != "All") %>% 
  filter(reef_extent_rel_region == max(reef_extent_rel_region)) %>% 
  select(reef_extent_rel_region) %>% 
  mutate(reef_extent_rel_region = round(reef_extent_rel_region, 1)) %>% pull()

min_reef_extent_nb <- data_reef_extent %>% 
  filter(subregion != "All") %>% 
  filter(reef_extent_abs == min(reef_extent_abs)) %>% 
  mutate(subregion = str_trim(gsub("[A-z]", "", subregion))) %>% select(subregion) %>% pull()

min_reef_extent_value_abs <- data_reef_extent %>% 
  filter(subregion != "All") %>% 
  filter(reef_extent_abs == min(reef_extent_abs)) %>% 
  select(reef_extent_abs) %>% 
  mutate(reef_extent_abs = format(round(reef_extent_abs, 0),
                                  big.mark = ",", scientific = FALSE)) %>% pull()

min_reef_extent_value_rel <- data_reef_extent %>% 
  filter(subregion != "All") %>% 
  filter(reef_extent_rel_region == min(reef_extent_rel_region)) %>% 
  select(reef_extent_rel_region) %>% 
  mutate(reef_extent_rel_region = round(reef_extent_rel_region, 1)) %>% pull()

```

The `r region_i` region is divided into `r nb_subregions` subregions (**Fig. 2.`r nb_chapter`.1**). Sub-region `r max_reef_extent_nb` hosts `r max_reef_extent_value_rel`% of coral reefs in the region (`r max_reef_extent_value_abs` km²) whereas sub-region `r min_reef_extent_nb` hosts only `r min_reef_extent_value_rel`% of coral reefs in the region (`r min_reef_extent_value_abs` km², **Supp. Table 1**).

```{r echo=FALSE, message=FALSE}

data_population <- read.csv("../../figs/08_text-gen/human_population.csv") %>% 
  filter(region == region_i)

pop_2020 <- data_population %>% filter(subregion == "All") %>% select(pop_2020) %>% 
  mutate(pop_2020 = format(round(pop_2020, 0), big.mark = ",", scientific = FALSE)) %>% pull()

pop_change_rel <- data_population %>% filter(subregion == "All") %>% select(pop_change_rel) %>% 
  mutate(pop_change_rel = round(pop_change_rel, 1)) %>% pull()

pop_change_abs <- data_population %>% filter(subregion == "All") %>% select(pop_change_abs) %>% 
  mutate(pop_change_abs = format(round(pop_change_abs/20, 0), big.mark = ",", scientific = FALSE)) %>% pull()

pop_change_abs_max <- data_population %>% filter(subregion != "All") %>% 
  filter(pop_change_abs == max(pop_change_abs)) %>% 
  mutate(subregion = str_trim(gsub("[A-z]", "", subregion))) %>% select(subregion) %>% pull()

pop_2020_max <- data_population %>% filter(subregion != "All") %>% 
  filter(pop_2020 == max(pop_2020)) %>% 
  select(subregion) %>% 
  mutate(subregion = parse_number(subregion)) %>% pull()

```

As of 2020, a total of `r pop_2020` inhabitants were living within 5 km of coral reefs. From 2000 to 2020, the region recorded a `r pop_change_rel`% increase in populations residing within 5 km of coral reefs, representing an average annual increase of `r pop_change_abs` inhabitants. The greatest increase in populations over this period was recorded in sub-region `r pop_change_abs_max`, although the highest number of inhabitants in 2020 was located in sub-region `r pop_2020_max` (**Supp. Table 2**).

</div>

\newpage

## Environmental pressures and disturbances

### Temperature regime

<div custom-style="explanations">
**[Number of words of the section: ~150]**
</div>

<div custom-style="text">

```{r echo=FALSE, message=FALSE}

data_sst <- read.csv("../../figs/08_text-gen/thermal_regime.csv") %>% 
  filter(region == region_i) %>% 
  filter(subregion == "All")

mean_sst <- data_sst %>% select(mean_sst) %>%
  mutate(mean_sst = round(mean_sst, 2)) %>% pull()

sst_increase <- data_sst %>% select(sst_increase) %>%
  mutate(sst_increase = round(sst_increase, 2)) %>% pull()

warming_rate <- data_sst %>% select(warming_rate) %>%
  mutate(warming_rate = round(warming_rate, 3)) %>% pull()

```

From 1985 to 2024, the long-term average of Sea Surface Temperature (SST) recorded over coral reefs within the `r region_i` region was `r mean_sst`°C (**Supp. Figure 1**). During this period, the SST over coral reefs in the region increased by `r sst_increase`°C, corresponding to an average warming rate of `r warming_rate`°C per year (**Supp. Table 3**).

\

```{r echo=FALSE, message=FALSE}

figure_path <- paste0("../../figs/03_part-2/fig-2/", str_to_lower(region_name), ".png")

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=100%}

**Figure 2.`r nb_chapter`.`r fig_counter`.** Annual mean Sea Surface Temperature (SST) anomalies (°C) over the coral reefs of `r region_i` from 1986 to 2024. SST anomalies were calculated based on the long-term average over the period 1986-2024. Positive values (red bars) indicate SST warmer than the long-term average, whereas negative values (blue bars) indicate SST cooler than the long-term average.

\newpage

From 1986 to 2024, periods of major heat stress occurred in ...

\

```{r echo=FALSE, message=FALSE}

figure_path <- paste0("../../figs/03_part-2/fig-3/", str_to_lower(region_name), ".png")

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=100%}

**Figure 2.`r nb_chapter`.`r fig_counter`.** Percentage of the coral reefs within the `r region_i` region experiencing heat stress levels per year, from 1986 to 2024. No stress (DHW = 0); Warning (possible bleaching, 0–4 DHW); Alert 1 (reef-wide bleaching, 4–8 DHW); Alert 2 (reef-wide bleaching and mortality of heat-sensitive corals, 8–12 DHW); Alert 3 (multi-species mortality, 12–16 DHW); Alert 4 (severe multi-species mortality, 16–20 DHW); and Alert 5 (near-complete mortality, ≥20 DHW).

</div>

\newpage

### Cyclones

<div custom-style="explanations">
**[Number of words of the section: ~150]**
</div>

<div custom-style="text">

```{r echo=FALSE, message=FALSE}

regional_cyclone_name <- case_when(region_i == "EAS" ~ "typhoons",
                                   region_i == "Caribbean" ~ "hurricanes",
                                   TRUE ~ "cyclones")

first_sentence <- case_when(region_i == "EAS" ~ "In the EAS region, cyclones are known as typhoons. ",
                            region_i == "Caribbean" ~ "In the Caribbean region, cyclones are known as hurricanes. ",
                            TRUE ~ "")

load("../../data/07_cyclones/02_cyclones_extracted.RData")

data_cyclones_nb <- data_cyclones %>%
  st_drop_geometry() %>% 
  mutate(region = str_squish(gsub("[[:digit:]]+", "", subregion))) %>% 
  filter(region == region_i) %>% 
  filter(saffir != 0) %>% 
  group_by(region, saffir) %>% 
  summarise(n = n_distinct(ts_id)) %>% 
  group_by(region) %>% 
  mutate(n_tot = sum(n)) %>% 
  ungroup() %>% 
  select(n_tot) %>% 
  distinct() %>% 
  pull()

data_cyclones_year <- data_cyclones %>%
  st_drop_geometry() %>% 
  mutate(region = str_squish(gsub("[[:digit:]]+", "", subregion))) %>% 
  filter(region == region_i) %>% 
  filter(saffir != 0) %>%
  mutate(year = year(time)) %>% 
  group_by(region, year) %>% 
  summarise(n = n_distinct(ts_id)) %>% 
  ungroup() %>% 
  complete(region, year = seq(1980, 2024, 1), fill = list(n = 0)) %>% 
  group_by(region) %>% 
  summarise(nb_mean = round(mean(n), 1),
            nb_sd = round(sd(n), 1))

data_cyclones_mean <- data_cyclones_year %>% select(nb_mean) %>% pull()

data_cyclones_sd <- data_cyclones_year %>% select(nb_sd) %>% pull()

data_cyclones_highest <- data_cyclones %>%
  st_drop_geometry() %>% 
  mutate(region = str_squish(gsub("[[:digit:]]+", "", subregion))) %>% 
  filter(region == region_i) %>% 
  filter(windspeed == max(windspeed, na.rm = TRUE))

data_cyclones_highest_name <- data_cyclones_highest %>% 
  mutate(ts_name = str_to_sentence(ts_name)) %>% 
  select(ts_name) %>% 
  pull()

data_cyclones_highest_date <- data_cyclones_highest %>% 
  mutate(date = paste0(str_to_sentence(month(time, label = TRUE, abbr = FALSE, locale="EN-us")),
                       " ", year(time))) %>% 
  select(date) %>% 
  pull()

data_cyclones_highest_speed <- data_cyclones_highest %>% 
  select(windspeed) %>% 
  mutate(windspeed = round(windspeed, 0)) %>% 
  pull()

data_cyclones_highest_dist <- data_cyclones_highest %>% 
  select(dist) %>% 
  mutate(dist = round(dist, 0)) %>% 
  pull()

```

`r first_sentence`From 1980 to 2025, a total of `r data_cyclones_nb` `r regional_cyclone_name` has passed within 100 km of coral reefs within the `r region_i` region, representing an average of `r data_cyclones_mean` ± `r data_cyclones_sd` SD `r regional_cyclone_name` per year (**Fig. 2.`r nb_chapter`.4**). The `r gsub('.{1}$', '', regional_cyclone_name)` with the highest sustained wind speed over the studied period was the `r gsub('.{1}$', '', regional_cyclone_name)` `r data_cyclones_highest_name`, in `r data_cyclones_highest_date`, which passed `r data_cyclones_highest_dist` km from a coral reef with sustained wind speed of `r data_cyclones_highest_speed` km.h^-1^ (**Fig. 2.`r nb_chapter`.5**). Although the 100 km threshold was selected, `r regional_cyclone_name` passing at a greater distance from a coral reef, not represented within this figure, may also have had an impact on coral reefs [1].

\

```{r echo=FALSE, message=FALSE}

figure_path <- paste0("../../figs/03_part-2/fig-4/", str_to_lower(region_name), ".png")

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=100%}

**Figure 2.`r nb_chapter`.`r fig_counter`.** Number of tropical storms that passed within 100 km of the coral reefs of the `r region_i` region from 1980 to 2025. Colors correspond to the cyclone’s Saffir-Simpson category along the entire cyclone’s track.

\

\

```{r echo=FALSE, message=FALSE}

figure_path <- paste0("../../figs/03_part-2/fig-5/", str_to_lower(region_name), ".png")

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=100%}

**Figure 2.`r nb_chapter`.`r fig_counter`.**  Maximum sustained wind speed of tropical storms passed within 100 km of a coral reef in the `r region_i` region between 1980 and 2025. Colors correspond to the cyclone’s Saffir-Simpson category along the entire cyclone’s track. However, the values of sustained wind speed are extracted from the nearest tropical storm position from a coral reef. For this reason, some sustained wind speed values are below the lower threshold of category 1 Saffir-Simpson scale (*i.e.* 119 km.h^-1^).

</div>

\newpage

### Other regional disturbances

<div custom-style="explanations">
**[Number of words of the section: ~200 (350 if no cyclones section)]**
</div>

\newpage

## Status of coral reefs in the region

<div custom-style="explanations">
**[Number of words of the section: ~1,500]**
</div>

\

```{r echo=FALSE, message=FALSE}

figure_path <- paste0("../../figs/03_part-2/fig-6/", str_to_lower(region_name), ".png")

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=100%}

<div custom-style="text">

**Figure 2.`r nb_chapter`.`r fig_counter`.** Modeled temporal trends of benthic percentage cover in the `r region_i` region from 1980 to 2024 for hard coral (A) and macroalgae (B). The bold line represents the estimated average benthic cover while darker and lighter ribbons represent 80% and 95% credible intervals, respectively.

</div>

\newpage

## Status of coral reefs in the sub-regions

<div custom-style="explanations">
**[Number of words of the section: ~700]**
</div>

\

```{r echo=FALSE, message=FALSE}

figure_path <- paste0("../../figs/03_part-2/fig-7/", str_to_lower(region_name), ".png")

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=100%}

<div custom-style="text">

**Figure 2.`r nb_chapter`.`r fig_counter`.** Modeled temporal trends of hard coral cover in `r region_i`’s sub-regions from 1980 to 2024. The bold line represents the estimated average benthic cover while darker and lighter ribbons represent 80% and 95% credible intervals, respectively.

</div>

\newpage

\

```{r echo=FALSE, message=FALSE}

figure_path <- paste0("../../figs/03_part-2/fig-8/", str_to_lower(region_name), ".png")

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=100%}

<div custom-style="text">

**Figure 2.`r nb_chapter`.`r fig_counter`.** Modeled temporal trends of macroalgae cover in `r region_i`’s sub-regions from 1980 to 2024. The bold line represents the estimated average benthic cover while darker and lighter ribbons represent 80% and 95% credible intervals, respectively.

</div>

\newpage

## Description of monitoring data

<div custom-style="explanations">
**[Number of words of the section: ~300]**
</div>

<div custom-style="text">

```{r echo=FALSE, message=FALSE}

data_monitoring <- read.csv("../../figs/08_text-gen/monitoring.csv") %>% 
  filter(region == region_i) %>% 
  filter(subregion == "All")

nb_datasets <- data_monitoring %>% select(nb_datasets) %>% pull()

nb_sites <- data_monitoring %>% select(nb_sites) %>% pull()

nb_surveys <- data_monitoring %>% select(nb_surveys) %>% pull()

first_year <- data_monitoring %>% select(first_year) %>% pull()

last_year <- data_monitoring %>% select(last_year) %>% pull()

surveys_90_perc <- data_monitoring %>% select(surveys_90_perc) %>% pull()

surveys_sites_value <- read.csv("../../figs/08_text-gen/monitoring.csv") %>% 
  filter(region == region_i) %>% 
  filter(subregion != "All") %>% 
  mutate(across(c(nb_surveys, nb_sites), ~as.numeric(str_replace_all(.x, ",", "")))) %>% 
  mutate(perc_surveys = nb_surveys*100/(sum(nb_surveys)),
         perc_sites = nb_sites*100/(sum(nb_sites)))
  
min_surveys_value <- surveys_sites_value %>% 
  filter(perc_surveys == min(perc_surveys)) %>% 
  select(perc_surveys) %>% 
  mutate(perc_surveys = format(round(perc_surveys, 1), big.mark = ",", scientific = FALSE)) %>% pull()

min_surveys_name <- surveys_sites_value %>% 
  filter(perc_surveys == min(perc_surveys)) %>% 
  select(subregion) %>% 
  mutate(subregion = parse_number(subregion)) %>% pull()

min_sites_value <- surveys_sites_value %>% 
  filter(perc_sites == min(perc_sites)) %>% 
  select(perc_sites) %>% 
  mutate(perc_sites = format(round(perc_sites, 1), big.mark = ",", scientific = FALSE)) %>% pull()

max_surveys_value <- surveys_sites_value %>% 
  filter(perc_surveys == max(perc_surveys)) %>% 
  select(perc_surveys) %>% 
  mutate(perc_surveys = format(round(perc_surveys, 1), big.mark = ",", scientific = FALSE)) %>% pull()

max_surveys_name <- surveys_sites_value %>% 
  filter(perc_surveys == max(perc_surveys)) %>% 
  select(subregion) %>% 
  mutate(subregion = parse_number(subregion)) %>% pull()

max_sites_value <- surveys_sites_value %>% 
  filter(perc_sites == max(perc_sites)) %>% 
  select(perc_sites) %>% 
  mutate(perc_sites = format(round(perc_sites, 1), big.mark = ",", scientific = FALSE)) %>% pull()

```

Temporal trends in the percentage cover of benthic categories from 1980 to 2024 were estimated using data from `r nb_datasets` datasets encompassing `r nb_sites` monitoring sites. These data were collected through `r nb_surveys` surveys conducted between `r first_year` and `r last_year`, with 90% of surveys taking place after `r surveys_90_perc` (**Supp. Figure 2**). Of all surveys carried out in the region between 1980 and 2024, `r max_surveys_value`% (representing `r max_sites_value`% of surveyed sites) were conducted in sub-region `r max_surveys_name`, whereas only `r min_surveys_value`% (`r min_sites_value`% of sites) were conducted in sub-region `r min_surveys_name` (**Supp. Table 4**).

\

```{r echo=FALSE, message=FALSE}

figure_path <- paste0("../../figs/03_part-2/fig-9/", str_to_lower(region_name), ".png")

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=100%}

**Figure 2.`r nb_chapter`.`r fig_counter`.** Spatio-temporal distribution of benthic cover monitoring sites across the `r region_i` region. Sites that were monitored for the longest period are represented at the top of the sites monitored for fewer years. The scale bar provides distance at the equator and may not accurately represent distance over the entire latitudinal range of the region. The borders and nomenclature represented on this map are not endorsed by the GCRMN.

</div>

\newpage

## References

<div custom-style="text">

[1] Puotinen, M., Drost, E., Lowe, R., Depczynski, M., Radford, B., Heyward, A., & Gilmour, J. (2020). Towards modelling the future risk of cyclone wave damage to the world's coral reefs. Global Change Biology, 26(8), 4302-4315.

</div>

\newpage

## Acknowledgments of contributions

<div custom-style="explanations">

GCRMN will use a modified version of (https://credit.niso.org/) to describe each contributor’s specific contribution to the GCRMN publication. Regional writing teams are requested to record the contributions of regional contributors to define the following three categories:

1) Data Contributors. Persons who participated in the acquisition of a dataset (or its management that has been used in the analysis of the report.

2) Regional Coordinators. Persons who coordinated the production of the writing of the regional chapter.

3) Co-authors. Persons making substitutional contributions to the production of the report, such as the conception or design of a monitoring programme; or the acquisition, analysis or interpretation of the monitoring data for the report; AND drafting of a chapter or reviewing it; AND final approval of the version to be published; AND agreement to be accountable for the accuracy or integrity of their contributions to the report.

For Co-Authors, please record the following information: Firstname, Surname, Email, ORCID (if available).

For all contributors, please complete the region-specific CRediT table here: [ADD LINK]. The information recorded by the regional teams and provided to the editorial team will be used to develop the appropriate CRediT statement. 

</div>
