---
title: "Supplementary Materials"
output:
  word_document:
    reference_docx: styles.docx
editor: source
---

```{r echo=FALSE, message=FALSE}

require(tidyverse)
require(readxl)
require(flextable)

fig_counter <- 1

tbl_counter <- 1

```

\

\newpage

<div custom-style="text">

**Supplementary Table `r tbl_counter`.** Title.

```{r echo=FALSE, message=FALSE}

table <- read_xlsx("../../figs/06_supp-mat/supp-tbl-1_reef-extent.xlsx")

flextable(table, cwidth = c(1.26, 1.26, 1.26, 1.26, 1.26)) %>%
  bold(part = "header") %>% 
  set_table_properties(opts_word = list(repeat_header = FALSE)) %>% 
  border(border = officer::fp_border())

tbl_counter <- tbl_counter + 1

```

</div>

\newpage

```{r echo=FALSE, message=FALSE}

figure_path <- "../../figs/06_supp-mat/surveys-year.png"

```

![](`r figure_path`){width=80%}

<div custom-style="text">

**Supplementary Figure `r fig_counter`.** Title.

</div>

\newpage

```{r echo=FALSE, message=FALSE}

figure_path <- "../../figs/06_supp-mat/surveys-depth.png"

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=80%}

<div custom-style="text">

**Supplementary Figure `r fig_counter`.** Title.

</div>

\newpage

```{r echo=FALSE, message=FALSE}

figure_path <- "../../figs/06_supp-mat/fig-X_a.png"

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=80%}

<div custom-style="text">

**Supplementary Figure `r fig_counter`.** Title.

</div>

\newpage

```{r echo=FALSE, message=FALSE}

figure_path <- "../../figs/06_supp-mat/fig-X_b.png"

fig_counter <- fig_counter + 1

```

![](`r figure_path`){width=80%}

<div custom-style="text">

**Supplementary Figure `r fig_counter`.** Title.

</div>

\newpage

<div custom-style="text">

**Supplementary Table `r tbl_counter`.** List of individual datasets used to estimate benthic cover trends in hard coral and macroalgae cover in the ten GCRMN regions.

```{r echo=FALSE, message=FALSE}

table <- read_xlsx("../../figs/06_supp-mat/tbl-1_datasetid-per-region.xlsx")

flextable(table, cwidth = c(1.6, 4.7)) %>%
  bold(part = "header") %>% 
  set_header_labels(region = "Region",
                    datasetID = "Dataset (datasetID)") %>% 
  set_table_properties(opts_word = list(repeat_header = FALSE)) %>% 
  border(border = officer::fp_border())

tbl_counter <- tbl_counter + 1

```

</div>
